---
title: "Prediction Assignment"
author: "Carlos Araya"
date: "1 de Julio de 2016"
output: html_document
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

1. Load the data.

We load the data pml-training.csv and pml-testing.csv.

```{r}
library(caret)
pml_training <- read.csv("/media/caraya/DATA/CURSOS/02-Datascience/08 - Practical_Machine_Learning/PML_Assignment/pml-training.csv", na.strings = c("", "NA", "#DIV/0!"))
pml_testing <- read.csv("/media/caraya/DATA/CURSOS/02-Datascience/08 - Practical_Machine_Learning/PML_Assignment/pml-testing.csv", na.strings = c("", "NA", "#DIV/0!"))
```
In this case, we found a lot of data with missing values. We proceed to eliminate every variable with missing values from the original database.
```{r}
# Mark the variables with miss values
isAnyMissing <- sapply(pml_testing, function (x) any(is.na(x) | x == ""))
table(isAnyMissing)
# Select the variables without missing values and its names contain belt, arm, dumbbell, forearm
isPredictor <- !isAnyMissing & grepl("belt|[^(fore)]arm|dumbbell|forearm", names(isAnyMissing))
predCandidates <- names(isAnyMissing)[isPredictor]
predCandidates
```
There are 100 variables with missing values and 52 variables for model.
```{r}
## Subset the primary dataset to include only the predictor candidates and the outcome variable, classe.
varToInclude <- c("classe", predCandidates)
pml_training <- pml_training[, varToInclude]
#dim(pml_training)
#names(pml_training)

## Make classe into a factor.
pml_training[,"classe"] <- as.factor(pml_training[,"classe"])

## Split the dataset into a 60% training and 40% probing dataset.

## Define train and test for cross validation
seed <- 24568
inTrain <- createDataPartition(y=pml_training$classe,
                               p=0.75,list=FALSE) ## Originary 75%
training_0 <- pml_training[inTrain,] # DTrain
testing_0 <- pml_training[-inTrain,] # DProbe
```

2. Transform the data

We centering and scaling the final dataset.
```{r}
library(data.table)

## Preprocess the prediction variables by Centering and Scaling.
X <- training_0[, predCandidates]
preProc <- preProcess(X)
preProc
XCS <- predict(preProc, X)
trainingCS <- data.table(data.frame(classe = training_0[, "classe"], XCS))

## Apply the Centering and Scaling to the probing dataset.
X <- testing_0[, predCandidates]
XCS <- predict(preProc, X)
testingCS <- data.table(data.frame(classe = testing_0[, "classe"], XCS))
```

3. Modeling

```{r}
## Set the control parameters.
ctrl <- trainControl(classProbs=TRUE,
                     savePredictions=TRUE,
                     allowParallel=TRUE)
## We comment the script for the time
## system.time(trainingModel <- train(classe ~ ., data=trainingCS, method="rf"))
load(file="/media/caraya/DATA/CURSOS/02-Datascience/08 - Practical_Machine_Learning/PML_Assignment/trainingModel.RData", verbose=TRUE)
```

The final model
```{r}
trainingModel
hat <- predict(trainingModel, trainingCS)
confusionMatrix(hat, training_0[, "classe"])

## Evaluate the model on the probing dataset
hat <- predict(trainingModel, testingCS)
confusionMatrix(hat, testing_0[, "classe"])

## Display the final model
varImp(trainingModel)
trainingModel$finalModel
```

3. Forecast
```{r}
pml_testingCS <- predict(preProc, pml_testing[, predCandidates])
hat <- predict(trainingModel, pml_testingCS)
pml_testing <- cbind(hat , pml_testing)
subset(pml_testing, select=names(pml_testing)[grep("belt|[^(fore)]arm|dumbbell|forearm"
                                                   , names(pml_testing), invert=TRUE)])
```

4. Others.
```{r}
## Save training model object for later.
## save(trainingModel, file="trainingModel.RData")
## Load the training model.
## Save Workspace for later
## save.image("/media/caraya/DATA/CURSOS/02-Datascience/08 - Practical_Machine_Learning/PML_Assignment/Prediction_Assignment.RData")
## load(file="trainingModel.RData", verbose=TRUE)
## Get predictions and evaluate.
```